Parallel Information Set Generation for Kriegspiel

### 2012-06-08 sbac/paper.tex

\documentclass[11pt]{article}
\usepackage{amsmath,amsfonts,latexsym,graphicx}
\usepackage{fullpage}
\usepackage{url,hyperref}
\usepackage{subfig}
\usepackage{tikz}
\usepackage{examplep}
\usepackage{moreverb}
\usepackage{listings}
\lstset{language=C,numbers=left,basicstyle=\footnotesize,numberstyle=\footnotesize}
\usetikzlibrary{arrows,decorations.pathmorphing,backgrounds,fit,shapes.geometric}
\usepgflibrary{shapes.geometric}
%\usepackage{ijcai09}

\setlength{\oddsidemargin}{0in}
\setlength{\topmargin}{0in}
\setlength{\textwidth}{6.5in}
\setlength{\textheight}{8.5in}

@@

\newcommand{\etal}{{\em et al.}\ }
\newcommand{\assign}{\leftarrow}
\newcommand{\eps}{\epsilon}

%\newtheorem{lemma}{Lemma}
%\newtheorem{theorem}[lemma]{Theorem}
\begin{document}

\title{Information Set Generation in Kriegspiel}
\author{Abhishek Gupta\qquad Mark Richards\qquad Osman Sarood\\University of Illinois at Urbana-Champaign\\CS598LVK:
Parallel Combinatorial Search} \maketitle

\begin{abstract}
Information Set Generation is the identification of the set of paths in an imperfect information game tree that are
consistent with a player's observations.  The ability to reason about the possible state of the world is crucial to the
performance of game playing agents.  In this work, we discuss the problem of information set generation in the context
of kriegspiel (partially observable chess).  We implement the algorithm on top of a general purpose combinatorial search
engine and discuss its performance under different load balancing strategies and with different grainsize parameters.
We discuss some variations in the search strategy and their effect on performance.  \end{abstract}

\section{Introduction}
In imperfect information games, players do not have access to full knowledge of the world. Examples of imperfect
information include hidden cards in Poker~\cite{billings02challenge} or Bridge~\cite{ginsberg96partition}, hidden tiles
in Scrabble~\cite{richards07opponent}, or hidden pieces in Stratego or Kriegspiel (partially observable
chess)~\cite{li94chess}. The game tree nodes that are indistinguishable to a player because they differ only in the
information that is hidden to the player by rule are called that player's {\em information set.}  The ability to
estimate the value of the possible states and to reason about the probability distribution over those states is crucial
to playing imperfect information games well.

The term {\em belief state} is sometimes used interchangeably with information set to refer to a probability
distribution over possible worlds.  The latter term comes from the game theory community and is preferred here.  A node
in a game tree denotes not only the current state of the game, but also uniquely defines a path from the initial state
or root node.  Thus, a game tree node implicitly encodes not only the current state of the game, but also all the
history of all decisions made by all players up to that point in the game.  Knowing one's information set means knowing
all possible game histories.

For many games, solving the information set generation problem is trivial.  For example, in a poker game, it is easy to
see that the unseen cards held by a player's opponents may be any permutation of the cards not seen by that player
(i.e., hole cards and any revealed community cards).  After the betting rounds, it would not be reasonable to assume
that each of these permutations of unseen cards is equally probable, as betting decisions made by the players up to that
point would be affected by the quality of those players' cards.  But the set of {\em possible} hands for all of the
opponents is easy to conceive and enumerate.

In the game of kriegspiel, information set generation is non-trivial.  The player knows the opponent's position with
certainty when the game begins, but after the initial move there may be varying levels of uncertainty about the location
of the opponent's pieces.  Unlike poker, it is not possible to simply permute all of the opponent's possible pieces over
all of the squares not occupied by the player's own pieces.  A configuration of pieces for the opponent is only valid if
it can be reached by a legal sequence of moves.  In the general case, finding the nodes in an information set is a
combinatorial search problem.

Information set generation is not the same problem as solving a game
tree, although it is arguably a necessary subroutine for quality game play in imperfect information games. The standard
game tree problem is to find an equilibrium (i.e., minimax) strategy for the game. This can be done using alpha-beta
pruning for perfect information games or linear programming for imperfect information games. Information set generation,
by contrast, is the problem of finding sequences of moves in the game tree that are consistent with a player's
observations.

\section{Background}
In this section, we describe the game of kriegspiel, and the particular variation of it that we have implemented.

Kriegspiel was invented by Henry Michael Temple in 1899.  Historically, the game has required three parties: two
competing players and one referee.  Each player sits with a chessboard that is partitioned off so as not be seen by the
others.  One player controls the white pieces; the other controls the black pieces.  The referee keeps track of both
sets of pieces.  Each player has a full set of chess pieces.  He may position the opponent's pieces however he wishes on
his own board (if at all).  The players alternate moves as in a regular chess game, but instead of announcing their
moves out loud, they write the source and destination square of their desired play son a piece of paper and pass them to
the referee, so that the move is not heard by the other player. The referee, who knows the location of all pieces,
checks the legality of the move.  If the move is legal, the referee makes the move on his own board and announces that a
legal move has been made and that it is the other player's turn to move.  If the move is illegal because it is blocked
by an opponent's piece or would place or leave the active player in check, the referee announces, "No."  There is no
penalty for attempting an illegal move.  The player continues to attempt moves until one finally succeeds.  If a player
has no legal moves (because of stalemate or checkmate), the referee announces the game's result.  With the advent of
computers, it is convenient to play the game over a network, with a computer acting as the referee.

@@

The referee announces when a player is in check.  (And again, this announcement is heard by both players.)  There is
some variation in what additional information is supplied in this case.  In one popular variant, the referee declares
that the player is in check "by rank," "by file," "by diagonal," or "by knight."  In the case of a diagonal attack, the
referee may announce whether the checking piece is along the long or short diagonal (from the king's perspective).  It
is possible to be in check by two pieces at the same time, in which case the direction of both attacks would be
declared.  In our implementation, for simplicity, the referee declares only that the player is in check, without
specifying the nature of the attack.

@@

Because there is no penalty for making illegal moves, it is often to a player's advantage to attempt to make moves which
are likely to be illegal, as hearing such declarations from the referee can provide important information to a player
about the location of the opponent's pieces.  In particular, since a diagonal move by a pawn is legal only the case
where such a move would capture an opponent's piece, it could potentially be profitable to attempt all or many pawn
captures on every turn.  In order to speed the game up, popular variants of the game require the referee to make some
kind of declaration at the beginning of a player's turn with respect to the pawn captures that are available.  This may
be as simple as declaring "Try" to indicate that at least one pawn capture is possible.  In our implementation, the
referee declares all possible pawn captures.

In the case of a pawn promotion, the player making the pawn advancement would secretly notify the referee as to which
piece (queen, rook, bishop, or knight) he would like.  To reduce the number of bits needed to specify a move, our
implementation assumes and allows only promotion to queen.

@@

\subsubsection{Special Moves}
Our implementation supports the capability for pawns to move two spaces on their first turn.  However, we did not
implement support for castling or {\em en passant} pawn capture.  (These moves require maintaining additional bits of
information at each state that keep track of whether pieces have moved and/or whether pieces moved on the previous
turn.)

\subsection{Example}
\begin{figure}
%\includegraphics[width=0.2\textwidth]{images/1W.png}
\begin{minipage}{\textwidth}
\centering
\subfloat[1. d4]{\label{fig:1W}\includegraphics[width=0.2\textwidth]{images/1W.png}}
\qquad
\subfloat[1. ... a5]{\label{fig:1B}\includegraphics[width=0.2\textwidth]{images/1B.png}}
\qquad
\subfloat[2. Bg5]{\label{fig:2W}\includegraphics[width=0.2\textwidth]{images/2W.png}}
\qquad
\subfloat[2. ... b6]{\label{fig:2B}\includegraphics[width=0.2\textwidth]{images/2B.png}}
\end{minipage}
\begin{minipage}{\textwidth}
\centering
\subfloat[3. Nc3]{\label{fig:3W}\includegraphics[width=0.2\textwidth]{images/3W.png}}
\qquad
\subfloat[3. ... c5]{\label{fig:3B}\includegraphics[width=0.2\textwidth]{images/3B.png}}
\qquad
\subfloat[4. d5]{\label{fig:4W}\includegraphics[width=0.2\textwidth]{images/4W.png}}
\qquad
\subfloat[4. ... Na6]{\label{fig:4B}\includegraphics[width=0.2\textwidth]{images/4B.png}}
\end{minipage}
\begin{minipage}{\textwidth}
\centering
\subfloat[5. d6!]{\label{fig:5W}\includegraphics[width=0.2\textwidth]{images/5W.png}}
\qquad
\subfloat[5. ... f6!]{\label{fig:5B}\includegraphics[width=0.2\textwidth]{images/5B.png}}
\qquad
\subfloat[6. e4]{\label{fig:6W}\includegraphics[width=0.2\textwidth]{images/6W.png}}
\qquad
\subfloat[6. ... fg]{\label{fig:6B}\includegraphics[width=0.2\textwidth]{images/6B.png}}
\end{minipage}
\begin{minipage}{\textwidth}
\centering
\subfloat[7. Bc4]{\label{fig:7W}\includegraphics[width=0.2\textwidth]{images/7W.png}}
\qquad
\subfloat[7. ... ed]{\label{fig:7B}\includegraphics[width=0.2\textwidth]{images/7B.png}}
\qquad
\subfloat[8 Qd5!]{\label{fig:8W}\includegraphics[width=0.2\textwidth]{images/8W.png}}
\qquad
\subfloat[8 ... Qc7?]{\label{fig:8B}\includegraphics[width=0.2\textwidth]{images/8B.png}}
\end{minipage}
\begin{minipage}{\textwidth}
\centering
\subfloat[9 Qf7+]{\label{fig:9W}\includegraphics[width=0.2\textwidth]{images/9W.png}}
\qquad
\subfloat[9 ... d8]{\label{fig:9B}\includegraphics[width=0.2\textwidth]{images/9B.png}}
\qquad
\subfloat[10 Qxf8++]{\label{fig:10W}\includegraphics[width=0.2\textwidth]{images/8W.png}}
\end{minipage}
\caption{Sequence of moves}
\label{sequence}
\end{figure}

\begin{figure}
\begin{verbatim}
1. d4 {(:)}
   a5 {(:)}
2. Bg5 {(:)}
   b6 {(:)}
3. Nc3 {(:Bd8)}
   c5 {(:)}
4. d5 {(:Bd8;Tc5)}
   Na6 {(:)}
5. d6 {(:)}
   f6 {(:Td6)}
6. e4 {(:)}
   fxg5 {(Xg5:Td6;Tg5)}
7. Bc4 {(:)}
   exd6 {(Xd6:Td6)}
8. Qd5 {(:)}
   Qc7 {(:)}
9. Qf7+ {(:)} 
   d8 {:g6,f7,e7}
10. Qxf8++ {(1-0:)}
\end{verbatim}
\caption{Listing of a kriegspiel game resulting in checkmate by white after 10 moves.}
\label{listing}
\end{figure}

The concept of information sets is perhaps best explained through an example.  Figures~\ref{listing} and~\ref{sequence}
show a full listing and the associated pictorial simulation of an actual kriegspiel game from ~\cite{li94chess}.  We
have adapted the notation used by Wolfe for Berkeley Kriegspiel.  The moves are numbered in pairs: a move by white on
one line followed by black's move on the next line.  The actual move is shown first followed by a list of referee
announcements.  Captures announcements are prefixed with an X and give the location of the captured piece.  Similarly,
pawn tries announcements are prefixed with T.  Attempted moves that were ultimately declared illegal are shown in a list
following the ':'.  For example at move 9, (Figure~\ref{fig:9W}--\ref{fig:9B}), black attempts to block a potential
threat along the diagonal by moving a pawn to f6, then attempts to escape or capture a threat at f7.  When both of these
attempts fail, black can infer that his king is being threatened by a protected bishop or queen at f7.  The attempt to
move to e7 is black's way of finding out whether the threatening piece is a queen or bishop.  When this move fails,
black knows it is white's queen and ultimately retreats to d8.  Note that the failed moves by white on turns 3 and 4 are
a result of being blocked by the pawn at e2.

Note that the listing in Figure~\ref{listing} corresponds to the referee's view of the game and would be useful for
post-game analysis or commentary.  During the game, players would not have access to the full information.  The
transcript from white's perspective is shown in Figure~\ref{filteredlisting}.  Note that black's actual moves have been
replaced with ``??'' and lists of black's illegal moves have been replaced with the {\em number} of illegal attempted
moves.  For example, on move 9, black attempted three illegal moves.
\begin{figure}
\begin{verbatim}
1. d4 {(:)}
   ?? {(:0)}
2. Bg5 {(:)}
   ?? {(:0)}
3. Nc3 {(:Bd8)}
   ??  {(:0)}
4. d5 {(:Bd8;Tc5)}
   ??  {(:0)}
5. d6 {(:)}
   ?? {(:0,Td6)}
6. e4 {(:)}
   fxg5 {(Xg5:Td6;Tg5)}
7. Bc4 {(:)}
   ?? {(Xd6:Td6)}
8. Qd5 {(:)}
   ?? {(:0)}
9. Qf7+ {(:)} 
   ?? {:0}
10. Qxf8++ {(1-0:)}
\end{verbatim}
\caption{Listing of a kriegspiel game resulting in checkmate by white after 10 moves.}
\label{filteredlisting}
\end{figure}

The listing of a sequence of moves from the perspective of one player lists all of that player's moves exactly and
includes the declarations from the referee.  We refer to such a list as a player's observations.  Given a list of
observations for a sequence of moves, an information set is the set of all possible {\em sequences of moves for both
players}, that are consistent with those observations.  Figure~\ref{abbrevoutput} shows the information set for black
after both player's have made five moves.  From this list of possibilities, black can infer that white must have a
bishop at g5 and a pawn at d6.  There are 25 possible sequences of moves that are consistent with black's observations
up to that point. Table~\ref{bothtimes} shows the size of the information sets for each player for each move in the game
and the amount of time to find the full set on a single processor.  Appendix B shows the full output of the program for
the same point in the game (five moves for each player) and illustrates all of the possible positions in the belief
state.

>>>>

\documentclass[times, 10pt,twocolumn]{article} 
\usepackage{latex8}
\usepackage{times}
\usepackage{url,hyperref}
\usepackage{subfig}
\usepackage{tikz}
\usepackage{examplep}
\usepackage{moreverb}
\usepackage{algorithm}
\usepackage{boxedminipage}
\usepackage[noend]{algpseudocode}
\usepackage{listings}
\usepackage{amsmath,amsfonts,latexsym,graphicx}
\lstset{language=C,numbers=left,basicstyle=\footnotesize,numberstyle=\footnotesize}
\usetikzlibrary{arrows,decorations.pathmorphing,backgrounds,fit,shapes.geometric}
\usepgflibrary{shapes.geometric}

%\documentclass[11pt]{article}
%\usepackage{fullpage}
%\usepackage{ijcai09}

%\setlength{\oddsidemargin}{0in}
%\setlength{\topmargin}{0in}
%\setlength{\textwidth}{6.5in}
%\setlength{\textheight}{8.5in}

@@

\newcommand{\etal}{{\em et al.}\ }
\newcommand{\assign}{\leftarrow}
\newcommand{\eps}{\epsilon}
%\documentstyle[times,art10,twocolumn,latex8]{article}

%------------------------------------------------------------------------- 
% take the % away on next line to produce the final camera-ready version 
\pagestyle{empty}

%------------------------------------------------------------------------- 
\begin{document}

\title{Parallel Information Set Generation for Kriegspiel 
       }

\author{Mark Richards \and Abhishek Gupta \and Osman Sarood \\
%Department of Computer Science\\ 
%University of Illinois\\ 
%Urbana, IL\\ mdrichar@illinois.edu\\
% For a paper whose authors are all at the same institution, 
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'', 
% just like the second author.
%\and
%Second Author\\
%Institution2\\
%First line of institution2 address\\ Second line of institution2 address\\ 
%SecondAuthor@institution2.com\\
}

\maketitle
\thispagestyle{empty}

%\newtheorem{lemma}{Lemma}
%\newtheorem{theorem}[lemma]{Theorem}
%\begin{document}

%\title{Information Set Generation in Kriegspiel}
%\author{Abhishek Gupta\qquad Mark Richards\qquad Osman Sarood\\University of Illinois at Urbana-Champaign\\CS598LVK:
%Parallel Combinatorial Search} \maketitle

\begin{abstract}
Information Set Generation is the identification of the set of paths in an imperfect information game tree that are
consistent with a player's observations.  The ability to reason about the possible game history is crucial to the
performance of game playing agents.  In this work, we discuss the problem of information set generation in the context
of kriegspiel (partially observable chess).  We implement the algorithm on top of a general purpose combinatorial search
engine and discuss its performance under different load balancing strategies and with different grainsize parameters.
Our results suggest that our algorithm can be efficiently parallelized.  
\end{abstract}

\Section{Introduction}
In imperfect information games, players do not have access to full knowledge of the world. Examples of imperfect
information include hidden cards in Poker~\cite{billings02challenge} or Bridge~\cite{ginsberg96partition}, hidden tiles
in Scrabble~\cite{richards07opponent}, or hidden pieces in Stratego or Kriegspiel (partially observable
chess)~\cite{li94chess}. The game tree nodes that are indistinguishable to a player because they differ only in the
information that is hidden to the player by rule are called that player's {\em information set.}  The ability to
estimate the value of the possible states and to reason about the probability distribution over those states is crucial
to playing imperfect information games well.

The term {\em belief state} is sometimes used interchangeably with information set to refer to a probability
distribution over possible worlds.  The latter term comes from the game theory community and is preferred here.  A node
in a game tree denotes not only the current state of the game, but also uniquely defines a path from the initial state
or root node.  Thus, a game tree node implicitly encodes not only the current state of the game but also the complete
history of all decisions made by all players up to that point in the game, including the outcome of any chance moves
such as dice rolls or card shuffling.  Knowing one's information set means knowing all possible game histories.

For many specific games, solving the information set generation problem is trivial.  For example, in a poker game, it is easy to
see that the unseen cards held by a player's opponents may be any permutation of the cards not seen by that player
(i.e., hole cards and any revealed community cards).  After the betting rounds, it would not be reasonable to assume
that each of these permutations of unseen cards is equally probable, as betting decisions made by the players up to that
point would be affected by the quality of those players' cards.  But the set of {\em possible} hands for all of the
opponents is easy to conceive and enumerate.

In the game of kriegspiel, information set generation is much more complicated.  The player knows the opponent's position with
certainty when the game begins, but after the initial move there may be varying levels of uncertainty about the location
of the opponent's pieces.  Unlike poker, it is not possible to simply permute all of the opponent's possible pieces over
all of the squares not occupied by the player's own pieces.  A configuration of pieces for the opponent is only valid if
it can be reached by a legal sequence of moves.  In the general case, finding the nodes in an information set is a
combinatorial search problem.

Information set generation is not the same problem as solving a game tree, although it is arguably a necessary
subroutine for quality game play in imperfect information games.  In this work, we describe an approach to generating
information sets for the game of kriegspiel.

\Section{Background}
In this section, we describe the game of kriegspiel, and the particular variation of it that we have implemented.

Kriegspiel was invented by Henry Michael Temple in 1899.  Historically, the game has required three parties: two
competing players and one referee.  Each player sits with a chessboard that is partitioned off so as not be seen by the
other.  One player controls the white pieces; the other controls the black pieces.  The referee has a board on which she
keeps track of both sets of pieces.  Each player has a full set of chess pieces.  He may position the opponent's pieces
however he wishes on his own board (if at all).  The players alternate moves as in a regular chess game, but instead of
announcing their moves out loud, they write the source and destination square of their desired plays on a piece of paper
and pass them to the referee, so that the requested move is not known to the other player. The referee, who knows the location of
all pieces, checks the legality of the move.  If the move is legal, the referee makes the move on his own board and
announces that a legal move has been made and that it is the other player's turn to move.  If the move is illegal
because it is blocked by an opponent's piece or would place or leave the active player in check, the referee announces,
``No.''  There is no penalty for attempting an illegal move.  The player continues to attempt moves until one finally
succeeds.  If a player has no legal moves (because of stalemate or checkmate), the referee announces the game's result.
With the advent of computers, it is convenient to play the game over a network, with a computer acting as the referee.

@@

The referee announces when a player is in check.  (And again, this announcement is heard by both players.)  There is
some variation in what additional information is supplied in this case.  In one popular variant, the referee declares
that the player is in check ``by rank,'' ``by file,'' ``by diagonal,'' or ``by knight.''  In the case of a diagonal attack, the
referee may announce whether the checking piece is along the long or short diagonal (from the king's perspective).  It
is possible to be in check by two pieces at the same time, in which case the direction of both attacks would be
declared.  In our implementation, for simplicity, the referee declares only that the player is in check, without
specifying the nature of the attack.

@@

Because there is no penalty for making illegal moves, it is often to a player's advantage to attempt to make moves which
are likely to be illegal, as hearing such declarations from the referee can provide important information to a player
about the location of the opponent's pieces.  In particular, since a diagonal move by a pawn is legal only the case
where such a move would capture an opponent's piece, it could potentially be profitable to attempt all or many pawn
captures on every turn.  In order to speed the game up, popular variants of the game require the referee to make some
kind of declaration at the beginning of a player's turn with respect to the pawn captures that are available.  This may
be as simple as declaring ``Try'' to indicate that at least one pawn capture is possible.  In our implementation, the
referee declares all potential pawn captures.

In the case of a pawn promotion, the player making the pawn advancement would secretly notify the referee as to which
piece (queen, rook, bishop, or knight) he would like.  To reduce the number of bits needed to specify a move, our
implementation allows only promotion to queen.

@@

\SubSection{Special Moves}
Our implementation supports the capability for pawns to move two spaces on their first turn.  However, we did not
implement support for castling or {\em en passant} pawn capture.  (These moves require maintaining additional bits of
information at each state that keep track of whether pieces have moved and/or whether pieces moved on the previous
turn.)

\SubSection{Example}
%\includegraphics[width=0.2\textwidth]{images/1W.png}
%\begin{minipage}{\textwidth}
%\centering
%\subfloat[1. d4]{\label{fig:1W}\includegraphics[width=0.2\textwidth]{images/1W.png}}
%\qquad
%\subfloat[1. ... a5]{\label{fig:1B}\includegraphics[width=0.2\textwidth]{images/1B.png}}
%\qquad
%\subfloat[2. Bg5]{\label{fig:2W}\includegraphics[width=0.2\textwidth]{images/2W.png}}
%\qquad
%\subfloat[2. ... b6]{\label{fig:2B}\includegraphics[width=0.2\textwidth]{images/2B.png}}
%\end{minipage}
%\begin{minipage}{\textwidth}
%\centering
%\subfloat[3. Nc3]{\label{fig:3W}\includegraphics[width=0.2\textwidth]{images/3W.png}}
%\qquad
%\subfloat[3. ... c5]{\label{fig:3B}\includegraphics[width=0.2\textwidth]{images/3B.png}}
%\qquad
%\subfloat[4. d5]{\label{fig:4W}\includegraphics[width=0.2\textwidth]{images/4W.png}}
%\qquad
%\subfloat[4. ... Na6]{\label{fig:4B}\includegraphics[width=0.2\textwidth]{images/4B.png}}
%\end{minipage}
%\begin{minipage}{\textwidth}
%\centering
%\subfloat[5. d6!]{\label{fig:5W}\includegraphics[width=0.2\textwidth]{images/5W.png}}
%\qquad
%\subfloat[5. ... f6!]{\label{fig:5B}\includegraphics[width=0.2\textwidth]{images/5B.png}}
%\qquad
%\subfloat[6. e4]{\label{fig:6W}\includegraphics[width=0.2\textwidth]{images/6W.png}}
%\qquad
%\subfloat[6. ... fg]{\label{fig:6B}\includegraphics[width=0.2\textwidth]{images/6B.png}}
%\end{minipage}
%\begin{minipage}{\textwidth}
%\centering
%\subfloat[7. Bc4]{\label{fig:7W}\includegraphics[width=0.2\textwidth]{images/7W.png}}
%\qquad
%\subfloat[7. ... ed]{\label{fig:7B}\includegraphics[width=0.2\textwidth]{images/7B.png}}
%\qquad
%\subfloat[8 Qd5!]{\label{fig:8W}\includegraphics[width=0.2\textwidth]{images/8W.png}}
%\qquad
%\subfloat[8 ... Qc7?]{\label{fig:8B}\includegraphics[width=0.2\textwidth]{images/8B.png}}
%\end{minipage}
%\begin{minipage}{\textwidth}
%\centering
%\subfloat[9 Qf7+]{\label{fig:9W}\includegraphics[width=0.2\textwidth]{images/9W.png}}
%\qquad
%\subfloat[9 ... d8]{\label{fig:9B}\includegraphics[width=0.2\textwidth]{images/9B.png}}
%\qquad
%\subfloat[10 Qxf8++]{\label{fig:10W}\includegraphics[width=0.2\textwidth]{images/8W.png}}
%\end{minipage}
\begin{figure}[ht]
\centering
\begin{minipage}[b]{0.5\linewidth}
\includegraphics[width=0.7\textwidth]{images/9B.png}
\caption{Actual Position}
\label{fig:figure1}
\end{minipage}
\hspace{0.5cm}
\centering
\begin{minipage}[b]{0.5\linewidth}
\includegraphics[width=0.7\textwidth]{images/OneView.png}
\caption{White's view}
\label{fig:figure2}
\end{minipage}
\end{figure}

%\begin{figure}
%\begin{verbatim}
%1. d4 {(:)}
%   a5 {(:)}
%2. Bg5 {(:)}
%   b6 {(:)}
%3. Nc3 {(:Bd8)}
%   c5 {(:)}
%4. d5 {(:Bd8;Tc5)}
%   Na6 {(:)}
%5. d6 {(:)}
%   f6 {(:Td6)}
%6. e4 {(:)}
%   fxg5 {(Xg5:Td6;Tg5)}
%7. Bc4 {(:)}
%   exd6 {(Xd6:Td6)}
%8. Qd5 {(:)}
%   Qc7 {(:)}
%9. Qf7+ {(:)} 
%   d8 {:g6,f7,e7}
%10. Qxf8++ {(1-0:)}
%\end{verbatim}
%\label{listing}
%\caption{Listing of a kriegspiel game resulting in checkmate by white after 10 moves.}
%\end{figure}

The concept of information sets is perhaps best explained through an example.  (See also Appendix A.)
Figures~\ref{listing} and~\ref{sequence} show a full listing and the associated pictorial progression of an actual
kriegspiel game from ~\cite{li94chess}.  We have adapted the notation used by Wolfe for Berkeley Kriegspiel.  The moves
are numbered in pairs: a move by white on one line followed by black's move on the next line.  The actual move is shown
first followed by a list of referee announcements.  Captures announcements are prefixed with an X and give the location
of the captured piece.  Similarly, pawn tries announcements are prefixed with T.  Attempted moves that were ultimately
declared illegal are shown in a list following the `:'.  For example at move 9, (Figure~\ref{fig:9W}--\ref{fig:9B}),
black attempts to block a potential threat along the diagonal by moving a pawn to f6, then attempts to escape or capture
a threat at f7.  When both of these attempts fail, black can infer that his king is being threatened by a protected
bishop or queen at f7.  The attempt to move to e7 is black's way of finding out whether the threatening piece is a queen
or bishop.  When this move fails, black knows it is white's queen and ultimately retreats to d8.  Note that the failed
moves by white on turns 3 and 4 are a result of being blocked by the pawn at e2.

The listing in Figure~\ref{listing} corresponds to the referee's view of the game and would be useful for
post-game analysis or commentary.  During the game, players would not have access to the full information.  The
transcript from white's perspective is shown in Figure~\ref{filteredlisting}.  Note that black's actual moves have been
replaced with ``??'' and lists of black's illegal moves have been replaced with the {\em number} of illegal attempted
moves.  For example, on move 9, black attempted three illegal moves.
\begin{figure}
\begin{verbatim}
1. d4 {(:)}
   ?? {(:0)}
2. Bg5 {(:)}
   ?? {(:0)}
3. Nc3 {(:Bd8)}
   ??  {(:0)}
4. d5 {(:Bd8;Tc5)}
   ??  {(:0)}
5. d6 {(:)}
   ?? {(:0,Td6)}
6. e4 {(:)}
   fxg5 {(Xg5:Td6;Tg5)}
7. Bc4 {(:)}
   ?? {(Xd6:Td6)}
8. Qd5 {(:)}
   ?? {(:0)}
9. Qf7+ {(:)} 
   ?? {:0}
10. Qxf8++ {(1-0:)}
\end{verbatim}
\caption{Listing of a kriegspiel game resulting in checkmate by white after 10 moves.}
\label{filteredlisting}
\end{figure}

The listing of a sequence of moves from the perspective of one player lists all of that player's moves exactly and
includes the declarations from the referee.  We refer to such a list as a player's observations.  Given a list of
observations for a sequence of moves, an information set is the set of all possible {\em sequences of moves for both
players}, that are consistent with those observations.  In other words, any sequence of moves that would have generated
the same sequence of observations for black is in black's information set.  Figure~\ref{abbrevoutput} shows the
information set for black after both player's have made five moves.  From this list of possibilities, black can infer
that white must have a bishop at g5 and a pawn at d6.  There are 25 possible sequences of moves that are consistent with
black's observations up to that point. Table~\ref{bothtimes} shows the size of the information sets for each player for
each move in the game and the amount of time to find the full set on a single processor.  Appendix B shows the full
output of the program for the same point in the game (five moves for each player) and illustrates all of the possible
positions in the belief state.

\begin{figure}
\begin{mytinylisting}
1. Pa2:a4 Pa7:a5 2. Pd2:d4 Pb7:b6 3. Bc1:g5 Pc7:c5 \\
4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
\\
1. Pd2:d3 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pd3:d4 Pc7:c5 \\
4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
\\
1. Pd2:d4 Pa7:a5 2. Pa2:a4 Pb7:b6 3. Bc1:g5 Pc7:c5 \\
4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\

%1. Pd2:d4 Pa7:a5 2. Bc1:f4 Pb7:b6 3. Bf4:g5 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pa2:a3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pa2:a4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pb2:b3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pc2:c3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pe2:e3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pe2:e4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pf2:f3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pf2:f4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pg2:g3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pg2:g4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ph2:h3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ph2:h4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:a3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:c3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:d3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:c1 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ke1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ng1:f3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ng1:h3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
\end{mytinylisting}
\caption{Abbreviated output from example game.  Generating information sets for black after five turns by each player.
From this, black can infer that there is definitely a white bishop at g5 and a white pawn at d6.}
\label{abbrevoutput}
\end{figure}
